---
title: "R Notebook"
output: html_notebook
---

# Final Project EDA 

```{r}
# Libraries
install.packages("RMariaDB")
library(RMariaDB)
library(DBI)
library(ggplot2)
```

```{r}
# Connect to database
db_host <- Sys.getenv("DB_READ_ENDPOINT")
db_user <- Sys.getenv("DB_READ_USER")
db_pw   <- Sys.getenv("DB_READ_PASSWORD")
db_port <- Sys.getenv("DB_READ_PORT")
db_name <- Sys.getenv("DB_READ_DB")
db_drv  <- RMariaDB::MariaDB()

con <- dbConnect(db_drv, user=db_user, password=db_pw, dbname=db_name, host=db_host, port=db_port)

dbListTables(con)
```
```{r}
# Take a look at the first few rows
query1 <- "
SELECT * 
FROM music
LIMIT 10

;
"

result1 <- dbGetQuery(con, query1)
result1
```
```{r}
# Look at relationship between mode, energy, and danceability
query2 <- "
SELECT speechiness, instrumentalness, energy
FROM music

;
"

result2 <- dbGetQuery(con, query2)
```

```{r}
ggplot(data = result2, mapping = aes(x = speechiness, y = instrumentalness, col = energy)) +
  geom_point(size = 1)
```
Most of the tracks displaying a high degree of speechiness have 0 instrumentalness, probably indicating audio books, podcasts, spoken word, etc. 

There are a few outliers that have both high speechiness and instrumentalness which is interesting because from the variable descriptions these seem mutually exclusive.

More energetic tracks clustered more around the extremes of instrumental and spechiness?


```{r}
# Disconnect from database to clean up connection
dbDisconnect(con)
```
