---
title: "R Notebook"
output: html_notebook
---

# Final Project EDA 

```{r}
# Libraries
install.packages("RMariaDB")
library(RMariaDB)
library(DBI)
```

```{r}
# Connect to database
db_host <- Sys.getenv("DB_READ_ENDPOINT")
db_user <- Sys.getenv("DB_READ_USER")
db_pw   <- Sys.getenv("DB_READ_PASSWORD")
db_port <- Sys.getenv("DB_READ_PORT")
db_name <- Sys.getenv("DB_READ_DB")
db_drv  <- RMariaDB::MariaDB()

con <- dbConnect(db_drv, user=db_user, password=db_pw, dbname=db_name, host=db_host, port=db_port)

dbListTables(con)
```
```{r}

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
