---
title: "Officials in NHL Games with Prolific Player Scoring"
subtitle: "MAS Stats 405 Spring 2021 Week 4 Assignment"
author: 
- Britney Brown, Harrison DiStefano, Greg Eastman
- Lisa Kaunitz, Tianyang Liu, Jeremy Weidner
date: "05/02/2021"
geometry: 'left=12mm, right=12mm, top=10mm, bottom=15mm'
fontsize: 12pt
output:
  pdf_document:
    latex_engine: xelatex
    keep_tex:  true
    toc: no
    toc_depth: '3'
header-includes:
 \usepackage{background}
 \usepackage{float}
 \backgroundsetup{
 scale=1,
 color=black,
 angle=0,
 pages=all,
 contents={\includegraphics[width=16.5cm,height=40cm]{nhl_action_02.png}}
 }
 
---

<style type="text/css">.main-container {  max-width: 2800px;  margin-left: auto;  margin-right: auto;}</style>
```{r SET UP, echo=FALSE, message=FALSE, eval=TRUE, results='hide', warning=FALSE}

# load libraries 
library(RMySQL)
library(knitr)
library(xtable)

options(xtable.comment = FALSE) #### prevents commenting when printing xtable latex

projpath <- getwd() # set working directory as project path

if(!exists("xdbsock")) {
    xdbsock <- ""
    cat("\n", 
        "Parameter 'xdbsock' not found, setting to empty string for general usage", 
        "\n")
}

drv <- dbDriver("MySQL") # create driver object

# connect to professor's database
xdbuser <- Sys.getenv("MAS405_AWS_DZES_DB_RO_USER")
xpw     <- Sys.getenv("MAS405_AWS_DZES_DB_RO_PW")
xdbname <- Sys.getenv("MAS405_AWS_DZES_DB_RO_DBNAME")
xdbhost <- Sys.getenv("MAS405_AWS_DZES_DB_RO_HOST")
xdbport <- as.integer( Sys.getenv("MAS405_AWS_DZES_DB_RO_PORT") )

con <- dbConnect(drv, unix.sock=xdbsock,
                 user=xdbuser, password=xpw, dbname=xdbname, host=xdbhost, port=xdbport)

# check that all NHL tables of interest exist in database connection
xbool_tableExists <- dbExistsTable(con, 
                    c("nhl_gameDate_1","nhl_goaliesDate_1",
                      "nhl_playersDate_1","nhl_teamDate_1"))
xbool_tableExists # if all TRUE, success!
```


```{r runQueryDate, echo=FALSE, eval=TRUE, results='hide', warning=FALSE}

##### get min and max date

qstr <- "SELECT MIN(date), MAX(date) FROM nhl_playersDate_1"

xx <- dbGetQuery(con, qstr)
xx

xMinDataDate <- xx[1, 1]
xMaxDataDate <- xx[1, 2]

xminDD <- format( strptime(xMinDataDate, "%Y%m%d"), "%m-%d-%Y")
xmaxDD <- format( strptime(xMaxDataDate, "%Y%m%d"), "%m-%d-%Y")

xpntsCO <- 3 # number of goals of interest
```

<style type="text/css">
#resultsT1 {
  width: 140% ;
}
#resultsOfficials1 {
  width: 50% ;
}
</style>

# Introduction

**In the NHL, are some officials more likely than others to have officiated a game with a prolific individual player performance?**

Utilizing regular season NHL game and player data spanning from `r xminDD` to `r xmaxDD`, we produce the frequencies of officials in games where a player scored `r xpntsCO + 1` or more goals.

# Methodology

To answer the achieve the above result we use two objects from the given database: "nhl_gameDate_1" and "nhl_playersDate_1". The first database object provides information on game date, visiting teams, home teams, and the names of the officials. The second provides information on a player's first name, last name, the team, and total goals scored in a game. By joining these two database objects together, and filtering for players that scored more than `r xpntsCO` goals, we provide an answer to the above question of interest. 

```{r runQuery1, echo=FALSE, eval=TRUE, results='hide', warning=FALSE}
qstr <- "SELECT COLUMN_NAME 
         FROM INFORMATION_SCHEMA.COLUMNS 
         WHERE TABLE_NAME='nhl_gameDate_1'"
xx <- dbGetQuery(con, qstr)
#xx

qstr <- "SELECT COLUMN_NAME 
         FROM INFORMATION_SCHEMA.COLUMNS 
         WHERE TABLE_NAME='nhl_playersDate_1'"
xx <- dbGetQuery(con, qstr)
#xx

####### process also described in Method section above #######
# select game table (contains info on game date, visiting and home teams, and officials)
# join with players table (contains info on a player's goals) where the dates match and:
## 1. the game's visiting team matches the players' team OR
## 2. the game's home team matches the player's team
#### this results in a joined table with information needed to answer the question 
#### (the names of officials for each game and the number of goals each player made)
# then filter to only show players that scored more than 3 goals

qstr <-
paste0(
"
SELECT game.date
		, players.firstName
        , players.lastName
        , players.timeOnIce
        , players.goals
        , CASE WHEN players.team = 'Montr\xe9al Canadiens' THEN 'Montreal Canadiens'
			else players.team end as team
		, CASE WHEN game.VT = 'Montr\xe9al Canadiens' THEN 'Montreal Canadiens'
			else game.VT end as VT
		, CASE WHEN game.HT = 'Montr\xe9al Canadiens' THEN 'Montreal Canadiens'
			else game.HT end as HT
		, game.officials
FROM nhl_gameDate_1 game
JOIN nhl_playersDate_1 players ON (game.date = players.date AND game.VT = players.team) 
						OR (game.date = players.date AND game.HT = players.team)
WHERE players.goals > ", xpntsCO
)
qstr

# output query
df_out_sub <- dbGetQuery(con, qstr)

class(df_out_sub) # data.frame

######## previous code when using SELECT * rather than selecting just the columns of interest ########
# colnames(df_out) # note that there are non-unique column names
# colnames(df_out) <- make.names(names=colnames(df_out), unique = TRUE) # make all column names unique
# 
# 
# # subset only the variables of interest
# df_out_sub <- df_out[ , c("date", "firstName", "lastName", "timeOnIce", "goals", "team", "VT", "HT", "officials")]

Date <- df_out_sub[, 1]
# strip the time into readable format
DateFormat <- format( strptime(Date, "%Y%m%d"), "%m-%d-%Y")
df_out_sub[, 1] <- DateFormat

nHSgames <- nrow(df_out_sub) # 53 games
```



# Results

## Table 1
The first table describes the basic information associated with all games in our data where a player scored more than `r xpntsCO` goals. This table includes the date of the game, the player's first and last name, team name, as well as their time on ice and total goals scored. On the right-hand side of the table, there is also information regarding visitng and home teams alongside the list of officials for that game. 

## Table 2

The second table shows the frequency of officials who officiated games where a player scored more than 3 goals. In this table we see that officials Matt MacPherson, Steve Kozari, Brad Meier, Dan Schachte, David Brisebois, Mark Wheler, and Tim Peel all officiated 5 or 6 games where at least one player was capable of scoring more than 3 goals in a single game. This seems to be rare occurrence since most officials have 2 or less occurrences officiating games with prolific player scoring. 


\newpage

```{r runQuery1result, echo=FALSE, eval=TRUE, results='asis', table.width=20, warning=FALSE}

library(kableExtra)

### kable(df_out_sub, format="html", table.attr = "style='width:40% ;'")

##### we can use opts_knit$get("rmarkdown.pandoc.to") to determine whether we'd 
##### rather use kable or xtable to render a table 

if( opts_knit$get("rmarkdown.pandoc.to") %in% "html" ) {
  kable(df_out_sub, format="html", table.attr = "style='width:100% ;'")
}

if( opts_knit$get("rmarkdown.pandoc.to") %in% "latex" ) {
  
  xcaption <- "NHL games where a Player Scored More than 3 Goals."
  

  xxtb <- kable(df_out_sub, format="latex", caption=xcaption)
  suppressWarnings(kable_styling(xxtb, latex_options=c("scale_down", "HOLD_position", "striped") ))
  

  #x.big <- suppressWarnings(xtable(df_out_sub, label='tabbig'))

  #caption(x.big) <- xcaption

  ## xx <- suppressWarnings(print(x.big, tabular.environment='longtable', include.rownames=FALSE, print.results=FALSE))

  # xx <- 
  #   suppressWarnings(
  #   print(
  #     x.big, 
  #     tabular.environment='tabular', 
  #     include.rownames=FALSE, 
  #     print.results=FALSE
  #     )
  #   )
  
  #print(x.big, tabular.environment='longtable',
  #      floating=FALSE,include.rownames=FALSE, 
  #      hline.after = c(-1,nrow(x.big)),
  #      add.to.row = list(pos = list(0),
  #      command = "\\hline \\endhead "), file = "output.tex")
  # yy <- strsplit(xx, "\n")[[1]]
  # 
  # yy <- c(yy[1], "\\tiny", yy[2:length(yy)])
  # 
  # writeLines(yy, "output.tex")
  
}
```

<!-- \input{output.tex} -->

\newpage

```{r runOfficials, echo=FALSE, eval=TRUE, results='show', warning=FALSE}

# unlist the column of officials (list of 4 people) by the semicolon and use gsub to separate  out their names
xofficials <- gsub("^\\s*|\\s*$", "", (unlist(strsplit(df_out_sub[ , "officials"], ";"))))

# create a frequency table of officials
xoff_df <- as.data.frame(table(xofficials))

# reorder by decreasing frequency
xoff_df <- xoff_df[ order(xoff_df[ , "Freq"], decreasing=TRUE), ]
```


```{r runTableOfficials, echo=FALSE, eval=TRUE, results='asis', warning=FALSE}


if( opts_knit$get("rmarkdown.pandoc.to") %in% "html" ) {
  
  nuoffs <- nrow(xoff_df)
  
  #kable( xoff_df, format="html", table.attr = "id=\"resultsOfficials1\""  )
  xxtb <- kable( xoff_df, format="html", table.attr = "style='width:40% ;'"  )
  xxtb
  #kable_styling(xxtb)
}

if( opts_knit$get("rmarkdown.pandoc.to") %in% "latex" ) {

  nuoffs <- nrow(xoff_df) # 90 occurances
  
  rownames(xoff_df) <- NULL # remove row names for table
  colnames(xoff_df) <- c("official", "number of games") # rename columns
  
  ###### "{ }\\" -- creates some vertical space (removed since we have a page break)
  xcaption2 <- paste0("Frequency of Officials who Officiated Games where a Player Scored more than ", xpntsCO, " Goals (See Table 1)")
  
  xxtb <- kable(xoff_df, format="latex", caption = xcaption2, booktabs = TRUE, longtable = TRUE)

  kable_styling(xxtb, latex_options= c("hold_position","striped"), latex_table_env="longtable", repeat_header_text = "\\textit{(continued)}", repeat_header_continued = TRUE, position = "center", font_size = 9)


  # 
  # x.big <- suppressWarnings(xtable(xoff_df))
  # 
  # caption(x.big) <-  xcaption2
  # 
  # ## xx <- suppressWarnings(print(x.big, tabular.environment='longtable', include.rownames=FALSE, print.results=FALSE))
  # 
  # xx <- 
  #   suppressWarnings(
  #     print(
  #       x.big, 
  #       size='\\fontsize{7pt}{7pt}\\selectfont',
  #       tabular.environment='longtable', 
  #       include.rownames=FALSE, 
  #       print.results=FALSE
  #       )
  #     )
  # 
  # 
  # #yy <- strsplit(xx, "\n")[[1]]
  # 
  # #yy <- c(yy[1], "\\tiny", yy[2:length(yy)])
  # yy <- xx
  # 
  # writeLines(yy, "output_2.tex")
  # 
  # 
  
}

```

<!-- \input{output_2.tex} -->



